The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-23 paper/paper.tex
paper: make projections implicit, coh2frame corrections
paper: corrections, better typesetting, deps it
paper: strip deps it, tweak color instead
paper: tweak deps colors
paper: fix sub-super-script in deps
paper: adding missing args for depsrestr
paper: replace remaining aligns with arrays

\newcommandx{\depsp}[8][2,3,4,5,6,7,8]{
  \ensuremath{\mathsf{\color{russian-green}{deps}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4]}^{\joinargs[#5][#6][#7]}}
}

\newcommandx{\depstypep}[8][2,3,4,5,6,7,8]{
  \ensuremath{\mathsf{\color{russian-green}{DEPS}}_{
    \joinargs[\mathsf{\color{indian-yellow}{#1}}][#2][#3][#4]}^{\joinargs[#5][#6][#7]}}
}

@@

\newcommandx{\deps}[2][1,2]{\depsp{}[][][][#1][#2][][]}
\newcommandx{\fulldeps}[1][1]{\depsp{full}[][][][#1][][][]}
\newcommandx{\depscohs}[2][1,2]{\depsp{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohs}[1][1]{\depsp{fullcohs}[][][][#1][][][]}
\newcommandx{\depscoht}[2][1,2]{\depsp{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscoht}[1][1]{\depsp{fullcoht}[][][][#1][][][]}
\newcommandx{\depstype}[2][1,2]{\depstypep{}[][][][#1][#2][][]}
\newcommandx{\fulldepstype}[1][1]{\depstypep{full}[][][][#1][][][]}
\newcommandx{\depscohstype}[2][1,2]{\depstypep{cohs}[][][][#1][#2][][]}
\newcommandx{\fulldepscohstype}[1][1]{\depstypep{fullcohs}[][][][#1][][][]}
\newcommandx{\depscohttype}[2][1,2]{\depstypep{coht}[][][][#1][#2][][]}
\newcommandx{\fulldepscohttype}[1][1]{\depstypep{fullcoht}[][][][#1][][][]}

@@

\begin{align*}
  \framep[n][p]      & : \U                                                                                                                 & \mbox{for all $p\leq n$}   \\
  \framep[n-1][p]    & : \U                                                                                                                 & \mbox{for all $p\leq n-1$} \\
  \framep[n-2][p]    & : \U                                                                                                                 & \mbox{for all $p\leq n-2$} \\
  \restrf[n][p]      & : \Pi q.\, \Pi \omega.\, \framep[n][p] \rightarrow \framep[n-1][p]                                                   & \mbox{for all $p\leq n$}   \\
  \restrf[n-1][p]    & : \Pi q.\, \Pi \omega.\, \framep[n-1][p] \rightarrow \framep[n-2][p]                                                 & \mbox{for all $p\leq n-1$} \\
  \cohf[n][p]        & : \Pi qr\epsilon\omega.\, \restrf[n-1][p][q][\epsilon] \circ \restrf[n][p][\omega][r] =                                                           \\
                     & \;\;\;\;\restrf[n-1][p][r][\omega] \circ \restrf[n][p][\epsilon][q+1]                                                & \mbox{for all $p\leq n$}   \\
  \mathsf{hypF0}     & : \framep[n][0]         = \unitpoint
  \\
  \mathsf{hypFS}     & : \framep[n][p+1]       = \Sigma d: \framep[n+1][p]. \Pi \omega.\, \painting[n][p] (\restrf[n+1][p](d))              & \mbox{for all $p\leq n-1$} \\
  \mathsf{hypPrevF0} & : \framep[n-1][0]                                                                                       = \unitpoint                              \\
  \mathsf{hypPrevFS} & : \framep[n-1][p+1] = \Sigma d: \framep[n][p]. \Pi \omega.\, \painting[n-1][p] (\restrf[n][p](d))                    & \mbox{for all $p\leq n-2$} \\
  \mathsf{hypRF0}    & : \restrf[n][0]        = \ldots \texttt{depending on hypF0}                                                                                       \\
  \mathsf{hypRFS}    & : \restrf[n][p+1]      = \ldots \texttt{depending on hypFS and hypPrevFS}                                            & \mbox{for all $p\leq n-1$} \\
  \ldots             & \;\;\;\;\texttt{fields for painting, restrf, cohp}
\end{align*}

@@

\begin{align*}
  \framep[n][p]   & :                                                                                                           \U                                                                                    & \mbox{for all $p\leq n$} \\
  \painting[n][p] & :                                                                                                           \framep[n][p] \rightarrow \U                                                          & \mbox{for all $p\leq n$} \\
  \restrf[n][p]   & :                                                                                                           \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]           & \mbox{for all $p\leq n$} \\
  \restrp[n][p]   & :                                                                                                           \Pi q \leq n-p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                       \\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                                                                                                        & \mbox{for all $p\leq n$} \\
  \cohf[n][p]     & :                                                                                                           \Pi qr\epsilon\omega.\, \restrf[n][p][q][\epsilon] \circ \restrf[n+1][p][\omega][r] =                            \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][q+1][\epsilon]                                                                                                                             & \mbox{for all $p\leq n$} \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,                                                                                                                                                                              \\
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                                                                                                      \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d)                                                                                       & \mbox{for all $p\leq n$} \\
\end{align*}

@@

So far, using this recursive definition of  $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\restrf[n][\lbrack 0,p-1 \rbrack]$, we are able to specify the three first component of the exected dependent structure:
\begin{align*}
  \framep[n][p]             & :                        \U                                                & \mbox{for all $p\leq n$} \\
  \painting[n][p]           & :                        \framep[n][p] \rightarrow \U                      & \mbox{for all $p\leq n$} \\
  \restrf[n][p]             & :                        \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p]
  \left(\begin{array}{l}
            \framep[n][\lbrack 0,p-1 \rbrack]   \\
            \painting[n][\lbrack 0,p-1 \rbrack] \\
            \restrf[n][\lbrack 0,p-1 \rbrack]   \\
          \end{array}\right)
  \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                                                                              \\
\end{align*}
which can appropriately be reformulated using sequences for appropriate definitions of $\frametype[n][\lbrack 0,p \rbrack]$, $\paintingtype[n][\lbrack 0,p \rbrack]$ and $\restrftype[n][\lbrack 0,p \rbrack]$, since the dependencies of $\restrf[n][p]$ into $\restrf[n][\lbrack 0,p-1 \rbrack]$ makes it a very dependent function~\citep{Hickey96}:
\begin{align*}
  \framep[n][\lbrack 0,p \rbrack]   & : \frametype[n][\lbrack 0,p \rbrack]                                     \\
  \painting[n][\lbrack 0,p \rbrack] & : \paintingtype[n][\lbrack 0,p \rbrack](\framep[n][\lbrack 0,p \rbrack]) \\
  \restrf[n][\lbrack 0,p \rbrack]   & : \restrftype[n][\lbrack 0,p \rbrack]
  \left(\begin{array}{l}
            \framep[n][\lbrack 0,p-1 \rbrack]   \\
            \painting[n][\lbrack 0,p-1 \rbrack] \\
            \restrf[n][\lbrack 0,p-1 \rbrack]   \\
          \end{array}\right)                                                                    \\
\end{align*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \paintingtype[n][\lbrack 0,p-1 \rbrack] & \framep[n][\lbrack 0,p \rbrack]   & :      & \U                                                                                                                   \\
    \paintingtype[n]                        & \unitpoint                        & \defeq & \unittype                                                                                                            \\
    \paintingtype[n][\lbrack 0,p \rbrack]   & (\framep[n][\lbrack 0,p \rbrack]) & \defeq & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack]) \times (\framep[n][p] \rightarrow \U) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \depstype[n][\lbrack 0,p-1 \rbrack]    &        & \defeq                               &
    \left(\begin{array}{lll}
              \framep[n][\lbrack 0,p-1 \rbrack]   & : & \frametype[n][\lbrack 0,p-1 \rbrack]    \\
              \painting[n][\lbrack 0,p-1 \rbrack] & : & \paintingtype[n][\lbrack 0,p-1 \rbrack]
              (\framep[n][\lbrack 0,p-1 \rbrack])                                               \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   & : & \restrftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \framep[n][\lbrack 0,p-1 \rbrack]   \\
                  \painting[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                         \\
            \end{array}\right)  \\
    \framep[n+1][\lbrack 0,p \rbrack]      &
    \deps[n][\lbrack 0,p-1 \rbrack]        & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\
    \framep[n+1][\lbrack 0,0 \rbrack]      &
    \unitpoint                             & \defeq & (\unitpoint,\unittype)                 \\
    \framep[n+1][\lbrack 0,p+1 \rbrack]    &
    \deps[n][\lbrack 0,p \rbrack]          & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack]
              (\deps[n][\lbrack 0,p-1 \rbrack]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][\lbrack 0,p-1 \rbrack]).
              \, \layer[n][p]
              (\deps[n][\lbrack 0,p \rbrack])
              (d)                                \\
            \end{array}\right)                                                 \\
    \restrftype[n][\lbrack 0,p-1 \rbrack]  &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
            \end{array}\right) & :      & \U                                                \\
    \restrftype[n]                         &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)                 & \defeq &
    \unittype                                                                                \\
    \restrftype[n][\lbrack 0,p \rbrack]    &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p \rbrack]   \\
              \painting[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)   & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][\lbrack 0,p-1 \rbrack]
      :\restrftype[n][\lbrack 0,p-1 \rbrack]
      \left(\begin{array}{l}
                \framep[n+1][\lbrack 0,p-1 \rbrack]   \\
                \painting[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).         \\
      \Pi q \leq n-p.\,\Pi \epsilon.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][\lbrack 0,p-1 \rbrack]   \\
                \painting[n][\lbrack 0,p-1 \rbrack] \\
                \restrf[n][\lbrack 0,p-1 \rbrack]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}                       \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \U                                     \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq & \framep[n+1][p]
    (\deps[n][\lbrack 0,p-1 \rbrack])
    \rightarrow \U                                                                           \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \layertype[n][p]
    (\deps[n][\lbrack 0,p \rbrack])
    \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq &
    d \mapsto \Pi\epsilon.\,\painting[n][p](\deps.\restrf[n][p][0][\epsilon](d))             \\
  \end{array}
\end{equation*}

Note that we write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the
projections of a $\deps$. Also, by abuse of notation, we write $\unitpoint$ for
the only inhabitant of $\depstype[n]$.

@@

We can now formally define $\painting[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:
\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][p]        &
    \fulldeps[n]             & :      & \paintingtype[n+1][p]             \\
    \painting[n+1][n+1]      &
    \fulldeps[n]             & \defeq & \fulldeps.E                       \\
    \painting[n+1][p \leq n] &
    \fulldeps[n]             & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][\lbrack 0,p \rbrack]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                             \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][\lbrack 0,p-1 \rbrack] &
    \fulldeps[n]                          & :      & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack])) \\
    \painting[n+1]                        &
    \fulldeps[n]                          & \defeq & \unittype                                                                                                                 \\
    \painting[n+1][\lbrack 0,p \rbrack]   &
    \fulldeps[n]
                                          & \defeq &
    (\painting[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                                                                    \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    \fulldeps[n]                          & :      & \U                                                         \\
    \restrptype[n]                        & E      & \defeq                                         & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times             \\
                                          &        &                                                & \left(
    \begin{array}{l}
        \Pi q \leq n-p.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                                 \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                     \\
  \end{array}
\end{equation*}

@@

  \fontsize{7.2}{9}\selectfont

@@

              \framep[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]) \\
              \painting[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps[n])                            \\

@@

              \framep[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps.\deps[n][\lbrack 0,p \rbrack]) \\
              \painting[n+1][\lbrack 0,p \rbrack](\depscohs.\fulldeps[n])                          \\

@@

             \fulldeps.\deps.\restrf[n][p][q][\epsilon](\restrf[n+1][p][r][\omega]

@@

             \fulldeps.\deps.\restrf[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon]

@@

             \layer[n][p](\fulldeps.\deps[n][\lbrack 0,p \rbrack])(\restrf[n][p]

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]) & :     \fulldepstype[n+1]                                                       \\
  \overline{\fulldeps}(\fulldepscohs[n]) & \defeq \left(\begin{array}{l}
                                                            \overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,n \rbrack]) \\
                                                            \fulldepscohs.E                                                   \\
                                                          \end{array}\right) \\
\end{align*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p-1 \rbrack] &
    \fulldepscohs[n]                    & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1]                        &
    \fulldepscohs[n]                    & \defeq & \unittype                                                                      \\
    \restrp[n+1][\lbrack 0,p \rbrack]   &
    \fulldepscohs[n]
                                        & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                     \\
  \end{array}
\end{equation*}

>>>>

\newcommandx{\depsmacro}[5][4,5]{
  \ensuremath{\mathsf{\color{strawberry}{#1}}_{
    \joinargs[\mathsf{\color{#3}{#2}}]}^{\joinargs[#4][#5]}}
}

@@

\newcommandx{\deps}[2][1,2]{\depsmacro{deps}{restr}{russian-green}[#1][#2]}
\newcommandx{\fulldeps}[1][1]{\depsmacro{deps}{fullrestr}{russian-green}[#1]}
\newcommandx{\depscohs}[2][1,2]{\depsmacro{deps}{coh}{chestnut}[#1][#2]}
\newcommandx{\fulldepscohs}[1][1]{\depsmacro{deps}{fullcoh}{chestnut}[#1]}
\newcommandx{\depscoht}[2][1,2]{\depsmacro{deps}{coht}{chestnut}[#1][#2]}
\newcommandx{\fulldepscoht}[1][1]{\depsmacro{deps}{fullcoht}{chestnut}[#1]}
\newcommandx{\depstype}[2][1,2]{\depsmacro{DEPS}{restr}{russian-green}[#1][#2]}
\newcommandx{\fulldepstype}[1][1]{\depsmacro{DEPS}{fullrestr}{russian-green}[#1]}
\newcommandx{\depscohstype}[2][1,2]{\depsmacro{DEPS}{coh}{chestnut}[#1][#2]}
\newcommandx{\fulldepscohstype}[1][1]{\depsmacro{DEPS}{fullcoh}{chestnut}[#1]}
\newcommandx{\depscohttype}[2][1,2]{\depsmacro{DEPS}{coht}{chestnut}[#1][#2]}
\newcommandx{\fulldepscohttype}[1][1]{\depsmacro{DEPS}{fullcoht}{chestnut}[#1]}

@@

\begin{equation*}
  \begin{array}{lclr}
    \framep[n][p]      & : & \U                                                                                                                 & \mbox{for all $p\leq n$}   \\
    \framep[n-1][p]    & : & \U                                                                                                                 & \mbox{for all $p\leq n-1$} \\
    \framep[n-2][p]    & : & \U                                                                                                                 & \mbox{for all $p\leq n-2$} \\
    \restrf[n][p]      & : & \Pi q.\, \Pi \omega.\, \framep[n][p] \rightarrow \framep[n-1][p]                                                   & \mbox{for all $p\leq n$}   \\
    \restrf[n-1][p]    & : & \Pi q.\, \Pi \omega.\, \framep[n-1][p] \rightarrow \framep[n-2][p]                                                 & \mbox{for all $p\leq n-1$} \\
    \cohf[n][p]        & : & \Pi qr\epsilon\omega.\, \restrf[n-1][p][q][\epsilon] \circ \restrf[n][p][\omega][r] =                                                           \\
                       &   & \restrf[n-1][p][r][\omega] \circ \restrf[n][p][\epsilon][q+1]                                                      & \mbox{for all $p\leq n$}   \\
    \mathsf{hypF0}     & : & \framep[n][0]         = \unitpoint
    \\
    \mathsf{hypFS}     & : & \framep[n][p+1]       = \Sigma d: \framep[n+1][p]. \Pi \omega.\, \painting[n][p] (\restrf[n+1][p](d))              & \mbox{for all $p\leq n-1$} \\
    \mathsf{hypPrevF0} & : & \framep[n-1][0]                                                                                       = \unitpoint                              \\
    \mathsf{hypPrevFS} & : & \framep[n-1][p+1] = \Sigma d: \framep[n][p]. \Pi \omega.\, \painting[n-1][p] (\restrf[n][p](d))                    & \mbox{for all $p\leq n-2$} \\
    \mathsf{hypRF0}    & : & \restrf[n][0]        = \ldots \texttt{depending on hypF0}                                                                                       \\
    \mathsf{hypRFS}    & : & \restrf[n][p+1]      = \ldots \texttt{depending on hypFS and hypPrevFS}                                            & \mbox{for all $p\leq n-1$} \\
    \ldots             &   & \texttt{fields for painting, restrf, cohp}
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{lclr}
    \framep[n][p]   & : & \U                                                                                                  & \mbox{for all $p\leq n$} \\
    \painting[n][p] & : & \framep[n][p] \rightarrow \U                                                                        & \mbox{for all $p\leq n$} \\
    \restrf[n][p]   & : & \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]                         & \mbox{for all $p\leq n$} \\
    \restrp[n][p]   & : & \Pi q \leq n-p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                                     \\
                    &   & \painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                  & \mbox{for all $p\leq n$} \\
    \cohf[n][p]     & : & \Pi qr\epsilon\omega.\, \restrf[n][p][q][\epsilon] \circ \restrf[n+1][p][\omega][r] =                                          \\
                    &   & \restrf[n][p][\omega][r] \circ \restrf[n+1][p][q+1][\epsilon]                                       & \mbox{for all $p\leq n$} \\
    \cohp[n][p]     & : & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,                                                                                \\
                    &   & \restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                \\
                    &   & \restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d) & \mbox{for all $p\leq n$} \\
  \end{array}
\end{equation*}

@@

So far, using this recursive definition of  $\framep[n][\lbrack 0,p-1 \rbrack]$ and $\restrf[n][\lbrack 0,p-1 \rbrack]$, we are able to specify the three first component of the exected dependent structure:
\begin{equation*}
  \begin{array}{lclr}
    \framep[n][p]             & :                        & \U                                                & \mbox{for all $p\leq n$} \\
    \painting[n][p]           & :                        & \framep[n][p] \rightarrow \U                      & \mbox{for all $p\leq n$} \\
    \restrf[n][p]             & :                        & \Pi q \leq n-p.\, \Pi \epsilon.\, \framep[n+1][p]
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   \\
            \end{array}\right)
    \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                                                                                \\
  \end{array}
\end{equation*}
which can appropriately be reformulated using sequences for appropriate definitions of $\frametype[n][\lbrack 0,p \rbrack]$, $\paintingtype[n][\lbrack 0,p \rbrack]$ and $\restrftype[n][\lbrack 0,p \rbrack]$, since the dependencies of $\restrf[n][p]$ into $\restrf[n][\lbrack 0,p-1 \rbrack]$ makes it a very dependent function~\citep{Hickey96}:
\begin{equation*}
  \begin{array}{lcl}
    \framep[n][\lbrack 0,p \rbrack]   & : & \frametype[n][\lbrack 0,p \rbrack]                                     \\
    \painting[n][\lbrack 0,p \rbrack] & : & \paintingtype[n][\lbrack 0,p \rbrack](\framep[n][\lbrack 0,p \rbrack]) \\
    \restrf[n][\lbrack 0,p \rbrack]   & : & \restrftype[n][\lbrack 0,p \rbrack]
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   \\
            \end{array}\right)                                                                      \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \paintingtype[n][\lbrack 0,p-1 \rbrack] & \framep[n][\lbrack 0,p \rbrack] & :      & \U                                                                                                                   \\
    \paintingtype[n]                        & \unitpoint                      & \defeq & \unittype                                                                                                            \\
    \paintingtype[n][\lbrack 0,p \rbrack]   & \framep[n][\lbrack 0,p \rbrack] & \defeq & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack]) \times (\framep[n][p] \rightarrow \U) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \depstype[n][\lbrack 0,p-1 \rbrack]    &        & \defeq                               &
    \left(\begin{array}{lll}
              \framep[n][\lbrack 0,p-1 \rbrack]   & : & \frametype[n][\lbrack 0,p-1 \rbrack]    \\
              \painting[n][\lbrack 0,p-1 \rbrack] & : & \paintingtype[n][\lbrack 0,p-1 \rbrack]
              (\framep[n][\lbrack 0,p-1 \rbrack])                                               \\
              \restrf[n][\lbrack 0,p-1 \rbrack]   & : & \restrftype[n][\lbrack 0,p-1 \rbrack]
              \left(\begin{array}{l}
                  \framep[n][\lbrack 0,p-1 \rbrack]   \\
                  \painting[n][\lbrack 0,p-1 \rbrack] \\
                \end{array}\right)                                         \\
            \end{array}\right)  \\
    \framep[n+1][\lbrack 0,p \rbrack]      &
    \deps[n][\lbrack 0,p-1 \rbrack]        & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\
    \framep[n+1][\lbrack 0,0 \rbrack]      &
    \unitpoint                             & \defeq & (\unitpoint,\unittype)                 \\
    \framep[n+1][\lbrack 0,p+1 \rbrack]    &
    \deps[n][\lbrack 0,p \rbrack]          & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][\lbrack 0,p \rbrack]
              (\deps[n][\lbrack 0,p-1 \rbrack]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][\lbrack 0,p-1 \rbrack]).
              \, \layer[n][p]
              (\deps[n][\lbrack 0,p \rbrack])
              (d)                                \\
            \end{array}\right)                                                 \\
    \restrftype[n][\lbrack 0,p-1 \rbrack]  &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p-1 \rbrack]   \\
              \painting[n][\lbrack 0,p-1 \rbrack] \\
            \end{array}\right) & :      & \U                                                \\
    \restrftype[n]                         &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)                 & \defeq &
    \unittype                                                                                \\
    \restrftype[n][\lbrack 0,p \rbrack]    &
    \left(\begin{array}{l}
              \framep[n][\lbrack 0,p \rbrack]   \\
              \painting[n][\lbrack 0,p \rbrack] \\
            \end{array}\right)   & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][\lbrack 0,p-1 \rbrack]
      :\restrftype[n][\lbrack 0,p-1 \rbrack]
      \left(\begin{array}{l}
                \framep[n+1][\lbrack 0,p-1 \rbrack]   \\
                \painting[n+1][\lbrack 0,p-1 \rbrack] \\
              \end{array}\right).         \\
      \Pi q \leq n-p.\,\Pi \epsilon.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][\lbrack 0,p-1 \rbrack]   \\
                \painting[n][\lbrack 0,p-1 \rbrack] \\
                \restrf[n][\lbrack 0,p-1 \rbrack]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}                       \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \U                                     \\
    \layertype[n][p]                       &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq & \framep[n+1][p]
    (\deps[n][\lbrack 0,p-1 \rbrack])
    \rightarrow \U                                                                           \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & :      & \layertype[n][p]
    (\deps[n][\lbrack 0,p \rbrack])
    \\
    \layer[n][p]                           &
    \deps[n][\lbrack 0,p \rbrack]
                                           & \defeq &
    d \mapsto \Pi\epsilon.\,\painting[n][p](\restrf[n][p][0][\epsilon](d))                   \\
  \end{array}
\end{equation*}

Note that we write $\framep$, $\painting$, $\restrf$ for the
projections of a $\deps$. Also, by abuse of notation, we write $\unitpoint$ for
the only inhabitant of $\depstype[n]$.

@@

We can now formally define $\painting[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:
\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][p]        &
    \fulldeps[n]             & :      & \paintingtype[n+1][p]   \\
    \painting[n+1][n+1]      &
    \fulldeps[n]             & \defeq & E                       \\
    \painting[n+1][p \leq n] &
    \fulldeps[n]             & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\deps[n][\lbrack 0,p \rbrack]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                   \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \painting[n+1][\lbrack 0,p-1 \rbrack] &
    \fulldeps[n]                          & :      & \paintingtype[n+1][\lbrack 0,p-1 \rbrack](\framep[n+1][\lbrack 0,p-1 \rbrack](\deps[n][\lbrack 0,p-1 \rbrack])) \\
    \painting[n+1]                        &
    \fulldeps[n]                          & \defeq & \unittype                                                                                                       \\
    \painting[n+1][\lbrack 0,p \rbrack]   &
    \fulldeps[n]
                                          & \defeq &
    (\painting[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                                                          \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    \fulldeps[n]                          & :      & \U                                                         \\
    \restrptype[n]                        & E      & \defeq                                         & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times             \\
                                          &        &                                                & \left(
    \begin{array}{l}
        \Pi q \leq n-p.\Pi \epsilon.\Pi d:\framep[n+1][p](\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                       \\ \painting[n][p](\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                     \\
  \end{array}
\end{equation*}

@@

@@

              \framep[n+1][\lbrack 0,p \rbrack](\deps[n][\lbrack 0,p-1 \rbrack])     \\
              \painting[n+1][\lbrack 0,p \rbrack](\fulldeps[n])                      \\

@@

              \framep[n+1][\lbrack 0,p \rbrack](\deps[n][\lbrack 0,p \rbrack]) \\
              \painting[n+1][\lbrack 0,p \rbrack](\fulldeps[n])                \\

@@

             \restrf[n][p][q][\epsilon](\restrf[n+1][p][r][\omega]

@@

             \restrf[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon]

@@

             \layer[n][p](\deps[n][\lbrack 0,p \rbrack])(\restrf[n][p]

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{equation*}
  \begin{array}{lcl}
    \overline{\fulldeps}(\fulldepscohs[n]) & :      & \fulldepstype[n+1]                                        \\
    \overline{\fulldeps}(\fulldepscohs[n]) & \defeq & \left(\begin{array}{l}
                                                                \overline{\deps}(\depscohs[n][\lbrack 0,n \rbrack]) \\
                                                                E                                                   \\
                                                              \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p-1 \rbrack] &
    \fulldepscohs[n]                    & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldeps[n])) \\
    \restrp[n+1]                        &
    \fulldepscohs[n]                    & \defeq & \unittype                                              \\
    \restrp[n+1][\lbrack 0,p \rbrack]   &
    \fulldepscohs[n]
                                        & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))             \\
  \end{array}
\end{equation*}
